Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-30 article.tex
with weight values for section A.2

This appendix is structured as follows: Section~\ref{sapp:signal} describes the
differences in terms of signal processing. Section~\ref{sapp:averaging} then
provides a detailed presentation of the averaging
approach. Section~\ref{sapp:connectome} presents an approach to retrieve edge
directionality. Finally, Section~\ref{sapp:results} presents some experimental
results to validate the different steps of our proposal.

@@

Filters $f_1$ and $f_2$ display similar performances and thus were given equal
weight. Additionally, this weight were chosen equal as the weight selected for the simplified method. In contrast, filters $f_3$
and $f_4$ turn out to be individually less competitive and were therefore given
smaller weights in the tuned averaging. Yet, as further shown in
Section~\ref{sapp:results}, combining all $4$ filters proves to slightly
improve performance with respect to using $f_1$ and $f_2$ only.

@@

This approach is based on the following
observation. The rise of fluorescence of a neuron indicates its
activation. If another neuron is activated after a slight delay, this
could be a consequence of the activation of the first neuron and
therefore indicates a directed link in the connectome from the first to
the second neuron.  Given this observation, we have computed for every
pair $(i,j)$ the term:
\begin{align}
s_{i,j} = \sum_{t=1}^{T - 1} \mathbb{1}((x_j^{t+1} - x_i^t) \in \left[\phi_1, \phi_2\right])
\end{align}
that could be interpreted as an image of the  number of times
that neuron $i$ activates neuron $j$. $\phi_1$ and $\phi_2$ are
parameters whose values have been chosen in our experiments equal to   
$0.2$ and $0.5$, respectively. Their role is to 
define when the difference between $x_j^{t+1}$  and $x_i^t$ can
indeed be assimilated to an event where neuron $i$ activates neuron
$j$.

@@

This has lead us to investigate other ways for exploiting the
information about directionality contained in the matrix $z$. One of
those ways that gave good performance was to use as association
matrix: 
\begin{align}
q_{i,j} = weight * p_{i,j} + (1-weight) * z_{i,j}
\end{align}
with  $weight$ chosen close to 1 ($weight=0.997$). Note that with
values for $weight$ close to 1,   matrix $q$ only slightly uses the
information about directivity contained in $z$ to modify the  partial
correlation matrix $p$. We tried smaller values for $weight$ but those
were leading to worse results.

@@

\paragraph{On the contribution of the directionality heuristic.}

Table~\ref{tab:dir} compares AUROC and AUPRC with and without prediction
of edge orientation. Both AUROC and AUPRC are (very slightly) improved by
directing the edges.

\begin{table}[ht]\label{tab:dir}
\caption{Performance on \textit{normal-1,2,3,4} of ``Full Method'' with and
  without prediction of edge orientation}
\label{tab:directivity}

@@

Results of several methods on these 6 datasets are provided in
Table~\ref{tab:results_appendix}. They confirm what we observed on the
\textit{normal} datasets. Average partial correlation and its tuned variant,
i.e.,``Full method'', clearly outperform other network inference methods on all
datasets. PC is close to GENIE3 and GTE but still slightly worse. GENIE3
performs better than GTE most of the times.

>>>>

This appendix is structured as follows: Section~\ref{sapp:signal} describes
the differences in terms of signal processing. Section~\ref{sapp:averaging}
then provides a detailed presentation of the averaging approach.
Section~\ref{sapp:connectome} presents an approach to correct the $p_{i,j}$
values so as to take into account the edge directionality. Finally,
Section~\ref{sapp:results} presents some experimental results to validate the
different steps of our proposal.

@@

Filters $f_1$ and $f_2$ display similar performances and thus were given similar
weights (i.e., resp. $1$ and $0.9$). Additionally, these weights were chosen equal as the weights selected for the simplified method. In contrast, filters $f_3$
and $f_4$ turn out to be individually less competitive and were therefore given
smaller weights in the tuned averaging (i.e., resp. $0.01$ and $0.7$). Yet, as further shown in
Section~\ref{sapp:results}, combining all $4$ filters proves to slightly
improve performance with respect to using $f_1$ and $f_2$ only.

@@

This approach is based on the following
observation. The rise of fluorescence of a neuron indicates its
activation. If another neuron is activated after a slight delay, this
could be a consequence of the activation of the first neuron and
therefore indicates a directed link in the connectome from the first to
the second neuron.  Given this observation, we have computed for every
pair $(i,j)$ the term:
\begin{align}
s_{i,j} = \sum_{t=1}^{T - 1} \mathbb{1}((x_j^{t+1} - x_i^t) \in \left[\phi_1, \phi_2\right])
\end{align}
that could be interpreted as an image of the  number of times
that neuron $i$ activates neuron $j$. $\phi_1$ and $\phi_2$ are
parameters whose values have been chosen in our experiments equal to   
$0.2$ and $0.5$, respectively. Their role is to 
define when the difference between $x_j^{t+1}$  and $x_i^t$ can
indeed be assimilated to an event for which neuron $i$ activates neuron
$j$.

@@

This has lead us to investigate other ways for exploiting the
information about directionality contained in the matrix $z$. One of
those ways that gave good performance was to use as association
matrix: 
\begin{align}
q_{i,j} = weight * p_{i,j} + (1-weight) * z_{i,j}
\label{eqn:qij}
\end{align}
with  $weight$ chosen close to 1 ($weight=0.997$). Note that with
values for $weight$ close to 1,   matrix $q$ only slightly uses the
information about directivity contained in $z$ to modify the  partial
correlation matrix $p$. We tried smaller values for $weight$ but those
were leading to worse results.

@@

\paragraph{On the interest of using matrix $q$ rather than $p$ to take into account directivity.}

Table~\ref{tab:dir} compares AUROC and AUPRC with or without correcting the $p_{i,j}$ values according to Equation \ref{eqn:qij}. Both AUROC and AUPRC are (very slightly) improved by using information about directivity.

\begin{table}[ht]\label{tab:dir}
\caption{Performance on \textit{normal-1,2,3,4} of ``Full Method'' with and
  without using information about directivity.}
\label{tab:directivity}

@@

Results of several methods on these 6 datasets are provided in
Table~\ref{tab:results_appendix}. They confirm what we observed on the
\textit{normal} datasets. Average partial correlation and its tuned variant,
i.e.,``Full method'', clearly outperform other network inference methods on all
datasets. PC is close to GENIE3 and GTE but still slightly worse. GENIE3
performs better than GTE most of the times. Note that the "Full method" reported in this table does not use Equation \ref{eqn:qij} to slightly correct the values of $p_{i,j}$  to take into account directivity.
